\pkg{cv}: An \proglang{R} Package for Cross-Validating Regression Models

### 2024-06-17 scratch/notes_john/JSS-article/JSS-article-3.Rmd
update draft JSS paper for new models()

We illustrate with the polynomial regression models of varying degree for the \code{Auto} data, beginning by fitting and saving the 10 models:
```{r polyomial-models}
mlist <- vector(10, mode="list")
for (p in 1:10) mlist[[p]] <- lm(mpg ~ poly(horsepower, p), data = Auto)
names(mlist) <- paste0("m.", 1:10)
mlist[2] # e.g., the quadratic fit
```
We then apply \code{cv()} to the list of 10 models (the \code{data} argument is required):
```{r polynomial-regression-CV}
# 10-fold CV
mlist <- do.call(models, mlist) # create "modList" object

@@

# LOO CV
cv.auto.loo <- cv(mlist, data=Auto, k="loo")
cv.auto.loo[2] # e.g., for quadratic model
```
The \code{models()} function takes an arbitrary number of regression models as its arguments, which are optionally named, to create a \code{"modList"} object. Because we generated the polynomial regression models in a named list, we conveniently employ \code{do.call()} to supply the models as arguments to \code{models()}. The names created for the list (e.g., \code{"m.2"}) are then used for the models. We can also invoke the \code{plot()} method for \code{"cvModList"} objects to compare the models (see Figure \ref{fig:polynomial-regression-CV-graph-2}):

>>>>

We illustrate with the polynomial regression models of varying degree for the \code{Auto} data, beginning by fitting and saving the 10 models:
```{r polyomial-models}
mlist <- vector(10, mode="list")
for (p in 1:10) mlist[[p]] <- lm(mpg ~ poly(horsepower, p), data = Auto)
names(mlist) <- paste0("m.", 1:10)
mlist[2] # e.g., the quadratic fit
```
We then apply \code{cv()} to the list of 10 models (the \code{data} argument is required):
```{r polynomial-regression-CV}
# 10-fold CV
mlist <- models(mlist) # create "modList" object

@@

# LOO CV
cv.auto.loo <- cv(mlist, data=Auto, k="loo")
cv.auto.loo[2] # e.g., for quadratic model
```
The \code{models()} function takes an arbitrary number of regression models as its arguments, which are optionally named, to create a \code{"modList"} object. Alternatively, as here, we can supply a pre-constructed, optionally named, list of models as the single argument to \code{models()}. To visualize the results, comparing the models, we invoke the \code{plot()} method for the \code{"cvModList"} object returned by \code{cv()} (see Figure \ref{fig:polynomial-regression-CV-graph-2}):

### 2024-06-17 scratch/notes_john/JSS-article/JSS-article-3.tex
update draft JSS paper for new models()

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 45649
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.13181
bias-adjusted cross-validation criterion = 19.12404
95% CI for bias-adjusted CV criterion = (15.69034, 22.55774)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> m.auto.glm <- glm(mpg ~ poly(horsepower, 2), data=Auto)
R> boot::cv.glm(Auto, m.auto.glm)$delta # MSE, biased-corrected MSE
\end{CodeInput}
\begin{CodeOutput}
[1] 19.24821 19.24787
\end{CodeOutput}
\begin{CodeInput}
R> microbenchmark::microbenchmark(
+   hatvalues = cv(m.auto, k="loo"),
+   Woodbury = cv(m.auto, k="loo", method="Woodbury"),
+   naive = cv(m.auto, k="loo", method="naive"),
+   cv.glm = boot::cv.glm(Auto, m.auto.glm),
+   times=10
+ )
\end{CodeInput}
\begin{CodeOutput}
Warning in microbenchmark::microbenchmark(hatvalues = cv(m.auto, k = "loo"), :
less accurate nanosecond times to avoid potential integer overflows
\end{CodeOutput}
\begin{CodeOutput}
Unit: microseconds
      expr        min         lq       mean     median         uq        max
 hatvalues    984.369   1182.604   1188.963   1202.345   1220.242   1293.468
  Woodbury  12366.543  12508.690  13408.099  12769.347  12815.124  16566.050
     naive 216467.126 217748.581 225187.170 219534.685 222515.077 276207.898
    cv.glm 373527.917 380937.683 393430.301 382341.584 389266.341 443234.190
 neval
    10
    10
    10
    10
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # 10-fold CV
R> mlist <- do.call(models, mlist) # create "modList" object
R> 
R> cv.auto.10 <- cv(mlist, data=Auto, seed=2120)
R> cv.auto.10[2] # e.g., for quadratic model
\end{CodeInput}
\begin{CodeOutput}

@@

The \code{models()} function takes an arbitrary number of regression
models as its arguments, which are optionally named, to create a
\code{"modList"} object. Because we generated the polynomial regression
models in a named list, we conveniently employ \code{do.call()} to
supply the models as arguments to \code{models()}. The names created for
the list (e.g., \code{"m.2"}) are then used for the models. We can also
invoke the \code{plot()} method for \code{"cvModList"} objects to
compare the models (see Figure
\ref{fig:polynomial-regression-CV-graph-2}):

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cv(hsb.lmer, seed=1575)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 1575
\end{CodeOutput}
\begin{CodeOutput}
Warning in checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv, :
Model failed to converge with max|grad| = 0.00587207 (tol = 0.002, component 1)
\end{CodeOutput}
\begin{CodeOutput}
boundary (singular) fit: see help('isSingular')
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
criterion: mse
cross-validation criterion = 37.44473
bias-adjusted cross-validation criterion = 37.33801
95% CI for bias-adjusted CV criterion = (36.28761, 38.38841)
full-sample criterion = 36.06767 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat) {
  if (!all(y %in% c(0, 1)))
    stop("response values not all 0 or 1")
  if (any(yhat < 0) ||
      any(yhat > 1))
    stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x145d6f3a0>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library("caret", quietly=TRUE, warn.conflicts=FALSE)
R> # reports complement of Bayes-rule criterion:
R> train(x = attrition[, c("JobSatisfaction", "Gender",
+                                "MonthlyIncome")],
+              y = attrition$Attrition,
+              method = "glm",
+              trControl = trainControl(method="LOOCV"))
\end{CodeInput}
\begin{CodeOutput}
Generalized Linear Model

@@

\begin{CodeChunk}
\begin{CodeInput}
R> all.equal(cv.attrition$`CV crit`,
+           cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="Woodbury")$`CV crit`,
+           )
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> all.equal(cv.attrition$`CV crit`,
+           cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="hatvalues")$`CV crit`
+           )
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> microbenchmark::microbenchmark(
+   rsample={
+     rs_obj$results <- map(rs_obj$splits, holdout_results, mod_form)
+     rs_obj$accuracy <- map_dbl(rs_obj$results,
+                                function(x) mean(x$correct))
+   },
+   caret=train(x = attrition[, c("JobSatisfaction", "Gender",
+                                 "MonthlyIncome")],
+               y = attrition$Attrition,
+               method = "glm",
+               trControl = trainControl(method="LOOCV")),
+   direct={
+     n <- nrow(attrition)
+     yhat <- numeric(n)
+     for (i in 1:n){
+       m <- update(mod.attrition, data=attrition[-i, ])
+       yhat[i] <- predict(m, newdata=attrition[i, ], type="response")
+     }
+     mean(((attrition$Attrition == "Yes") - round(yhat))^2)
+   },
+   cv.exact=cv(mod.attrition, k="loo", criterion=BayesRule),
+   cv.wood=cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="Woodbury"),
+   cv.hatvalues=cv(mod.attrition, k="loo", criterion=BayesRule,
+                   method="hatvalues"),
+   times=10
+ )
\end{CodeInput}
\begin{CodeOutput}
Warning in microbenchmark::microbenchmark(rsample = {: less accurate nanosecond
times to avoid potential integer overflows
\end{CodeOutput}
\begin{CodeOutput}
Unit: milliseconds
         expr         min          lq        mean      median          uq
      rsample 5741.137299 5867.248666 6155.761099 6095.937708 6218.031915
        caret 6805.241869 6931.865408 7135.256404 7052.951569 7377.336681
       direct 5262.786773 5312.612884 5455.994681 5388.428711 5475.565416
     cv.exact 4740.642384 4771.993936 4870.961647 4847.941434 4981.140799
      cv.wood  104.161361  106.036578  112.172855  113.107500  115.602534
 cv.hatvalues    1.262554    1.462716    1.497295    1.520198    1.569685
         max neval
 7162.266671    10
 7648.302729    10
 6045.941344    10
 5054.087753    10
  120.361527    10
    1.666896    10
\end{CodeOutput}
\end{CodeChunk}

>>>>

Residual standard error: 4.374 on 389 degrees of freedom
Multiple R-squared:  0.6876,    Adjusted R-squared:  0.686 
F-statistic:   428 on 2 and 389 DF,  p-value: < 2.2e-16
\end{CodeOutput}
\begin{CodeInput}
R> cv(m.auto, confint=TRUE)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 969978
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
method: Woodbury
criterion: mse
cross-validation criterion = 19.18252
bias-adjusted cross-validation criterion = 19.17193
95% CI for bias-adjusted CV criterion = (15.72359, 22.62027)
full-sample criterion = 18.98477 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> m.auto.glm <- glm(mpg ~ poly(horsepower, 2), data=Auto)
R> boot::cv.glm(Auto, m.auto.glm)$delta # MSE, biased-corrected MSE
\end{CodeInput}
\begin{CodeOutput}
[1] 19.24821 19.24787
\end{CodeOutput}
\begin{CodeInput}
R> microbenchmark::microbenchmark(
+   hatvalues = cv(m.auto, k="loo"),
+   Woodbury = cv(m.auto, k="loo", method="Woodbury"),
+   naive = cv(m.auto, k="loo", method="naive"),
+   cv.glm = boot::cv.glm(Auto, m.auto.glm),
+   times=10
+ )
\end{CodeInput}
\begin{CodeOutput}
Warning in microbenchmark::microbenchmark(hatvalues = cv(m.auto, k = "loo"), :
less accurate nanosecond times to avoid potential integer overflows
\end{CodeOutput}
\begin{CodeOutput}
Unit: microseconds
      expr        min         lq       mean     median         uq        max
 hatvalues    996.956   1181.292   1496.971   1216.777   1233.157   4243.664
  Woodbury  12189.054  12270.726  13187.437  12580.645  12664.121  16369.455
     naive 216220.101 218010.571 231464.303 220980.816 223550.942 280828.393
    cv.glm 381634.027 383670.456 404475.689 390533.979 441471.559 442930.011
 neval
    10
    10
    10
    10
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # 10-fold CV
R> mlist <- models(mlist) # create "modList" object
R> 
R> cv.auto.10 <- cv(mlist, data=Auto, seed=2120)
R> cv.auto.10[2] # e.g., for quadratic model
\end{CodeInput}
\begin{CodeOutput}

@@

The \code{models()} function takes an arbitrary number of regression
models as its arguments, which are optionally named, to create a
\code{"modList"} object. Alternatively, as here, we can supply a
pre-constructed, optionally named, list of models as the single argument
to \code{models()}. To visualize the results, comparing the models, we
invoke the \code{plot()} method for the \code{"cvModList"} object
returned by \code{cv()} (see Figure
\ref{fig:polynomial-regression-CV-graph-2}):

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cv(hsb.lmer, seed=1575)
\end{CodeInput}
\begin{CodeOutput}
R RNG seed set to 1575
\end{CodeOutput}
\begin{CodeOutput}
Warning in checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv, :
Model failed to converge with max|grad| = 0.00587131 (tol = 0.002, component 1)
\end{CodeOutput}
\begin{CodeOutput}
boundary (singular) fit: see help('isSingular')
\end{CodeOutput}
\begin{CodeOutput}
10-Fold Cross Validation
criterion: mse
cross-validation criterion = 37.44473
bias-adjusted cross-validation criterion = 37.33801
95% CI for bias-adjusted CV criterion = (36.28761, 38.38841)
full-sample criterion = 36.06767 
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> BayesRule
\end{CodeInput}
\begin{CodeOutput}
function(y, yhat) {
  if (!all(y %in% c(0, 1)))
    stop("response values not all 0 or 1")
  if (any(yhat < 0) ||
      any(yhat > 1))
    stop("fitted values outside of interval [0, 1]")
  yhat <- round(yhat)
  result <- mean(y != yhat) # proportion in error
  attr(result, "casewise loss") <- "y != round(yhat)"
  result
}
<bytecode: 0x13b56c108>
<environment: namespace:cv>
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> library("caret", quietly=TRUE, warn.conflicts=FALSE)
\end{CodeInput}
\begin{CodeOutput}

Attaching package: 'ggplot2'
\end{CodeOutput}
\begin{CodeOutput}
The following object is masked _by_ '.GlobalEnv':

    mpg
\end{CodeOutput}
\begin{CodeOutput}
The following object is masked from 'package:latticeExtra':

    layer
\end{CodeOutput}
\begin{CodeInput}
R> # reports complement of Bayes-rule criterion:
R> train(x = attrition[, c("JobSatisfaction", "Gender",
+                                "MonthlyIncome")],
+              y = attrition$Attrition,
+              method = "glm",
+              trControl = trainControl(method="LOOCV"))
\end{CodeInput}
\begin{CodeOutput}
Generalized Linear Model

@@

\begin{CodeChunk}
\begin{CodeInput}
R> all.equal(cv.attrition$`CV crit`,
+           cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="Woodbury")$`CV crit`,
+           )
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> all.equal(cv.attrition$`CV crit`,
+           cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="hatvalues")$`CV crit`
+           )
\end{CodeInput}
\begin{CodeOutput}
[1] TRUE
\end{CodeOutput}
\begin{CodeInput}
R> microbenchmark::microbenchmark(
+   rsample={
+     rs_obj$results <- map(rs_obj$splits, holdout_results, mod_form)
+     rs_obj$accuracy <- map_dbl(rs_obj$results,
+                                function(x) mean(x$correct))
+   },
+   caret=train(x = attrition[, c("JobSatisfaction", "Gender",
+                                 "MonthlyIncome")],
+               y = attrition$Attrition,
+               method = "glm",
+               trControl = trainControl(method="LOOCV")),
+   direct={
+     n <- nrow(attrition)
+     yhat <- numeric(n)
+     for (i in 1:n){
+       m <- update(mod.attrition, data=attrition[-i, ])
+       yhat[i] <- predict(m, newdata=attrition[i, ], type="response")
+     }
+     mean(((attrition$Attrition == "Yes") - round(yhat))^2)
+   },
+   cv.exact=cv(mod.attrition, k="loo", criterion=BayesRule),
+   cv.wood=cv(mod.attrition, k="loo", criterion=BayesRule,
+              method="Woodbury"),
+   cv.hatvalues=cv(mod.attrition, k="loo", criterion=BayesRule,
+                   method="hatvalues"),
+   times=10
+ )
\end{CodeInput}
\begin{CodeOutput}
Unit: milliseconds
         expr         min          lq        mean     median          uq
      rsample 5892.991705 5997.040644 6197.075663 6180.82761 6298.898798
        caret 6863.187210 6952.637853 7008.026442 7024.60940 7072.257280
       direct 5483.414538 5488.340073 5603.901283 5564.68712 5655.604370
     cv.exact 4829.717139 4885.941915 5049.808394 5021.14425 5155.600842
      cv.wood  104.694812  110.656458  114.844686  117.06384  119.173552
 cv.hatvalues    1.281824    1.513761    1.656785    1.59406    1.668454
         max neval
 6871.248999    10
 7095.635316    10
 5972.080828    10
 5486.199381    10
  121.229702    10
    2.391448    10
\end{CodeOutput}
\end{CodeChunk}
